---
title: "HW1"
author: "Cody Carroll"
date: "Fall 2023"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

# Question 1 
a.  I'm interested in applying R and DS to my field which is psychology. 
b. $y = 2\pi t$ 
c. In this next chunk I am going to use the function `print()`


```{r, eval = TRUE, echo = TRUE}
print("Hello World")
```

The options  `eval = TRUE` and `echo = TRUE` tell the .Rmd doc to run the code chunk and print the output in the .Rmd file. 


Next, I'll do the same as above except I will tell R **not** to print out my input:
```{r, echo = FALSE, eval = TRUE}
print("Hello World")
```

Next let's generate 1000 observations from a N(0,1) distribution and print the histogram of the data with 100 bins.

```{r, echo = TRUE, eval = TRUE}
data <- rnorm(n = 1000, mean = 0, sd = 1)
hist(data, n = 100)
```

Finally, to convert this entire .Rmd doc to a .pdf or .html file, click "Knit>Knit to pdf" or "Knit>Knit to html" at the top of the source file window.